# Master Script

## Introduction

A key way to streamline workflows in data management is by using a master script that runs other multiple scripts. This approach centralizes execution, ensuring consistency and efficiency across tasks. A master script acts as the central hub that coordinates and runs other scripts, automating various tasks and ensuring consistency across processes.

## Advantages

1.  **Efficiency**: It saves time by automating repetitive tasks, reducing the need for manual intervention.

2.  **Consistency**: Ensures that all tasks are performed in the same sequence and manner each time, minimizing errors.

3.  **Time saving**: Instead of running each data cleaning script individually, the master script can execute them all at once.

4.  **Scalability**: You can increase the number of scripts by adding them to the master script.

By integrating all essential steps into a single execution file, a master script optimizes data management, making the entire process more reliable and seamless.

## Structure of The Script

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, ft.align = "left")
options(gtsummary.tbl_summary.percent_fun = 
          function(x) sprintf(x * 100, fmt='%#.1f'))
```

Run the script form the first to the last.

```{r}

#| label: Structure of script

#--- Import data 

source("scripts/01-data-import.R")

# ---- Clean data

source("scripts/02-data-cleaning.R")
```
